---
title: "Lab1"
author: "Pasha Perevedentsev"
date: "10 Apr 2015"
output: pdf_document
---
# Preparation
- Load graph at http://leonidzhukov.net/hse/2015/sna
- Start RStudio

# Task 1

```{r}
library('igraph')
library('ggplot2')
```

Now, we are going to generate power distribution syntetically. The functions below are for generating random variables distributed according to power law and creating histogramm based on generated data.

```{r, warning=FALSE}
generate_dist = function(xmin, xmax, alpha, size) {
  #uniformly distributed variables
  r = runif(size)
  #generate power distributed variables from uniformly ones
  return((xmin^(-alpha+1)+r*(xmax^(-alpha+1)-xmin^(-alpha+1)))^(1/(-alpha+1)))
}
#commands bellow separate our date into bins with predefined number of bins
cut_bins = function(x, breaks){
  return(table(cut(x, breaks=breaks, labels=1:breaks)))
}
```

The next step is generating series from power laws having different $\alpha$:

```{r, warning=FALSE}
alphas = seq(2.1, 3.5 ,by=0.15)
num_of_groups = length(alphas)
n = 20
#sapply applies function to every element of list with alphas
data = sapply(alphas, function(alpha)
  cut_bins(generate_dist(1, 40, alpha, 10000), 40)[0:n])
#the result is matrix where each column correspond
#to bins generated from sets of variables with different alphas
```

Now, $data$ keeps our series. Let's look at them:

```{r, warning=FALSE}
head(data, 3)
```

Every column contain generated series for particular $\alpha$. To plot this data we have to do some transformations.

```{r, warning=FALSE}
data = data.frame(data)
colnames(data) = sapply(alphas, function(x) toString(x))
data = stack(data)
data$i = rep(1:n, num_of_groups)
data$ind = factor(data$ind, levels = data$ind)
```

At the first three rows we have transformed our data to stacked form. Following lines add additional information to our dataset. Let's look at our data now:

```{r, warning=FALSE}
head(data)
```

More informative representation is a graph:

```{r, warning=FALSE}
p = ggplot(data, aes(x=i, y=values, group=ind, colour=ind))
p + geom_line() + xlab('Amount of connections') + ylab('Number of vertexes')
``` 

It can be (hardly) seen that growth of $\alpha$ leads to smoothing of the curve. In other words, bigger 
$\alpha$ corresponds to more equal distribution of edges.

Assume we have data about degrees of graph's nodes. What the ways of calculating $\alpha$ parameter?
Recall that probability density function (PDF) has following form:
$$p(x)=Cx^{-\alpha}$$

Logarithm both sides: $$\log{p(x)}=\log(C) - \alpha \log{x}$$
Calculate $\log{p(x)}$ and $\log{x}$ from observed data and fit linear regression to find $\alpha$:

```{r, warning=FALSE}
r = c()
#As you remember, we have stacked data.
#We are going to interate over sets of rows having
#different values of indicator variable
for (i in unique(data$ind)){
  #Take rows with particular indicator and value bigger then zero
  prepared_data = data[data$ind == i & data$values > 0,]
  #Regress and store results to lm.fit
  lm.fit = lm(data = prepared_data, formula = log(values) ~ log(i))
  #Extract coefficient from regression and store it into r
  r = c(r, -coef(summary(lm.fit))[2,1])
}
plot(alphas, alphas-r, main='Alpha error against aplha (PDF Regression)')
```

Another way is to use cumulative distribution function (CDF): $cdf(x) = \int_{-\inf}^{x} p(x) dx = 1 - Cx^{1-\alpha}$:

```{r, warning=FALSE}
r = c()
for (i in unique(data$ind)){
  prepared_data = data[data$ind == i & data$values > 0,]
  t = c()
  #Following loop restores original form of data
  for (j in 1:nrow(prepared_data)) {
    t = c(t, rep(prepared_data[j, 'i'], prepared_data[j, 'values']))
  }
  prepared_data = t  
  #Then we create data.frame containing values of CDF function
  df = data.frame(c(1:n), c(sapply(1:n, ecdf(prepared_data))))
  colnames(df) = c('i', 'v')
  df$v = 1 - df$v
  #Remove singularities
  df = df[df$v > 0, ]
  lm.fit = lm(data = df, formula = log(v) ~ log(i))
  r = c(r, 1-coef(summary(lm.fit))[2,1] )
}
plot(alphas, alphas-r,main='Alpha error against aplha (CDF regression)')
```

Or we can just use embedded function of igraph:

```{r, warning=FALSE}
r = c()
for (i in unique(data$ind)){
  prepared_data = data[data$ind == i & data$values > 0,]
  t = c()
  for (j in 1:nrow(prepared_data)) {
    t = c(t, rep(prepared_data[j, 'i'], prepared_data[j, 'values']))
  }
  prepared_data = t
  r = c(r, power.law.fit(prepared_data, 1, impelementation = "plfit")$alpha)
}
plot(alphas, alphas-r, main='Alpha error against aplha (Maximal LL)')
```

The final part of the task is to calculate parameters and goodness-of-fit for several networks:
- link to network 1
- link to network 2

Hints:

`g = read.graph ...`

`degree(g)`

`power.law.fit`

`fit$alpha`

`fit$KS.p`