---
title: "Magolego SNA - Lab 3"
author: "Yulia Dodonova, Dmitry Petrov, Andrey Shestakov"
date: "04/17/2015"
output: 
  html_document:
    toc: true
---
```{r}
library('igraph')
library('ggplot2')
```

## Erdos-Reni random graph model

### Generate random graphs

There is a special function in {igraph} to generate random graphs according to the Erdos-Renyi model.
You can specify either the probability for drawing an edge between two arbitrary nodes or the number of edges in the graph:

```{r}
n=20
p=0.2
m=(n*(n-1)/2)*p
er11 <- erdos.renyi.game(n, p, type="gnp")
er12 <- erdos.renyi.game(n, m, type="gnm")
op <- par(mfrow = c(1, 2))
plot(er11, layout=layout.circle(er11), main="gnp")
plot(er12, layout=layout.circle(er12), main="gnm")
par(op)
print(length(E((er11))))
print(length(E((er12))))
```

Now we will vary the probability parameter and see how our graph changes:

```{r}
n=100
p1=0.01
p2=0.04
p3=0.1
er21 <- erdos.renyi.game(n, p1, type="gnp")
er22 <- erdos.renyi.game(n, p2, type="gnp")
er23 <- erdos.renyi.game(n, p3, type="gnp")
op <- par(mfrow = c(1, 3))
plot(er21, layout=layout.circle(er21), main="p=0.01")
plot(er22, layout=layout.circle(er22), main="p=0.04")
plot(er23, layout=layout.circle(er23), main="p=0.1")
par(op)
```

### Degree distribution

Now it's your turn - can you compute node degrees for these graphs and print summary statistics?

Hints:
`your_degrees = degree(your_graph)`

```{r}
der1 <- degree(er21)
summary(der1)
```

Now we will add degree distributions to the previous plots:

```{r}
der21 <- degree(er21)
der22 <- degree(er22)
der23 <- degree(er23)
op <- par(mfrow = c(2, 3))
plot(er21, layout=layout.circle(er21), main="p=0.01")
plot(er22, layout=layout.circle(er22), main="p=0.04")
plot(er23, layout=layout.circle(er23), main="p=0.1")
hist(der21, col=rgb(0,0,1,.4), xlim=c(0,20), xlab="degree",ylab="freq", main="p=0.01")
hist(der22, col=rgb(1,0,0,.4), xlim=c(0,20), xlab="degree",ylab="freq", main="p=0.04")
hist(der23, col=rgb(0,1,0,.4), xlim=c(0,20), xlab="degree",ylab="freq", main="p=0.1")
par(op)
```


### Phase transition and gigantic connected component

Now, let's explore gigantic connected components in random graphs with different probabilities:

```{r}
n1=10
n2=30
#n3=200
er311 <- erdos.renyi.game(n1, 0.5/n1, type="gnp")
er312 <- erdos.renyi.game(n1, 1.1/n1, type="gnp")
er313 <- erdos.renyi.game(n1, 4/n1, type="gnp")
er321 <- erdos.renyi.game(n2, 0.5/n2, type="gnp")
er322 <- erdos.renyi.game(n2, 1.1/n2, type="gnp")
er323 <- erdos.renyi.game(n2, 4/n2, type="gnp")

op <- par(mfrow = c(2, 3))
plot(er311, layout=layout.fruchterman.reingold(er311), main="n=10, np=0.5")
plot(er312, layout=layout.fruchterman.reingold(er312), main="n=10, np=1")
plot(er313, layout=layout.fruchterman.reingold(er313), main="n=10, np=4")
plot(er321, layout=layout.fruchterman.reingold(er321), main="n=30, np=0.5")
plot(er322, layout=layout.fruchterman.reingold(er322), main="n=30, np=1")
plot(er323, layout=layout.fruchterman.reingold(er323), main="n=30, np=4")
par(op)
```

How does the size of the gigantic connected component depend on the probability? Remember from the lecture?

```{r}
gcc=c()
cc=c()
probability=c()
for (i in 1:70){
  prob=0.0002*i
  probability=c(probability, prob)
  er <- erdos.renyi.game(500, prob, type="gnp")
  comp <- clusters(er)
  largest <- max(comp$csize)
  gcc=c(gcc, largest)
  clust <- transitivity(er, type="global")
  cc=c(cc, clust)
}

par(mfrow = c(1,1))
plot(probability*500, gcc, pch=19, col=26, xlab="np", main="Size of the gigantic connected component")

log(500)
```

### Trying to put ER model on real-networks

Load [Internet](http://www-personal.umich.edu/~mejn/netdata/as-22july06.zip) network in R.
Plot its degree distribution in log-log scale.

```{r, warning=FALSE}
g <- read.graph(file = 'as-22july06.gml', format = 'gml')
d <- degree.distribution(g)
# 
plot(d,log = 'xy', xlab = 'Degree d', ylab = 'p(d)', main = 'Degree distribution in log-log scale ')
```

Calculate average degree if the network and put a line for poisson distribution `dpois()` on the same plot `lines()`
```{r}
## Your code is here
```

Not even close! \
To finish with it identife the regime of this network as if it was Random Network. Determine the size of GCC and convice yourself that it conflicts with ER theory.
```{r}
## Your code is here
```

## Barabasi-Albert model. Preferential attachement
### Generate Barabasi-Albert graphs

Again, we can use a function in igraph

```{r}
b1=barabasi.game(30, power = 1, m=3, directed=F)
b2=barabasi.game(30, power = 1, m=10, directed=F)

op <- par(mfrow = c(1, 2))
plot(b1, layout=layout.fruchterman.reingold(b1), main="3 new edges")
plot(b2, layout=layout.fruchterman.reingold(b2), main="10 new edges")
par(op)
```

### Degree distribution

Now we will look at degree distributions for the Barabasi-Albert graphs. We will plot degree distributions for our small graphs and for a larger graph:

```{r}
b3=barabasi.game(1000, power = 1, m=10, directed=F)

db1 <- degree(b1)
db2 <- degree(b2)
db3 <- degree(b3)
op <- par(mfrow = c(1, 3))
hist(db1, breaks=10, col=rgb(0,0,1,.4), xlab="degree",ylab="freq", main="n=30, m=3")
hist(db2, breaks=10, col=rgb(1,0,0,.4),xlab="degree",ylab="freq", main="n=30, m=10")
hist(db3, breaks=50, col=rgb(1,0,0,.4),xlab="degree",ylab="freq", main="n=1000, m=10")
par(op)
```

Can you estimate $\alpha$ for the third graph? Remember how to do it?

Hints:

`power.law.fit`

`fit$alpha`

```{r}
fit <- power.law.fit(degree(b3))
fit$alpha
```

### Dinamical change

Now, we will look at how clustering coefficient, average path length and node degrees change with time.

```{r}
avpathlen=c()
clustcoef=c()
deg30=c()
deg70=c()
ba=barabasi.game(20, power = 1, m=3, directed=F)
m=15
for (i in 21:300){
  ba=add.vertices(ba, 1)
  neib=sample(seq(1,vcount(ba)), m, prob = degree(ba,normalized = T))
  new_edges=c()
  for (n in neib){
    new_edges=c(new_edges, c(i, n))
  }
  ba=add.edges(ba, new_edges)
  avpathlen=c(avpathlen, average.path.length(ba))
  clustcoef=c(clustcoef, transitivity(ba, type="global"))
  if (i>30){
    deg30=c(deg30, degree(ba, 30))}
  if (i>70){
    deg70=c(deg70, degree(ba, 70))}
  }
```

Now we will plot average path length and clustering coefficient against the number of nodes in a graph:

```{r}
xaxis=c(21:300)
op <- par(mfrow = c(1, 2))
plot(xaxis, avpathlen, pch=19, col=26, xlab="time", ylab="Average path length", main="Average path length")
plot(xaxis, clustcoef, pch=19, col=49, xlab="time", ylab="Clustering coefficient", main="Clustering coefficient")
par(op)
```

And now we will look at node degrees:

```{r}
plot(c(31:300), deg30, pch=19, col=35, ylim=c(0, max(deg30)), xlab="time", ylab="Node degree", main="Node degree")
points(c(71:300), deg70, pch=19, col=84)
legend("topleft", c("i=30", "i=70"), fill=c(35, 84))
```

## Small world model. Watts-Strogats model

Yet again, we will use a function in igraph:

```{r}
lat <- graph.lattice(length=4, dim=2, nei=1)
ws1 <- watts.strogatz.game(dim=2, size=4, nei=1, p=0.1, loops = FALSE, multiple = FALSE)
ws2 <- watts.strogatz.game(dim=2, size=4, nei=1, p=0.4, loops = FALSE, multiple = FALSE)
ws3 <- watts.strogatz.game(dim=2, size=4, nei=1, p=0.9, loops = FALSE, multiple = FALSE)

op <- par(mfrow = c(2, 2))
plot(lat, layout=layout.circle(lat), main="Lattice")
plot(ws1, layout=layout.circle(ws1), main="Rewiring probability 0.1")
plot(ws2, layout=layout.circle(ws2), main="Rewiring probability 0.4")
plot(ws3, layout=layout.circle(ws3), main="Rewiring probability 0.9")
par(op)
```

Now it's your turn again. Can you compute clustering coefficients and average path lengths for our small world graphs?

Hints:

`transitivity(your_graph, type="global")`

`average.path.length(your_graph)`

Can you plot degree distribution for these graphs?
