---
title: "Social Network Analysis <br/> Home Assignment 3"
date: due date - 05.06.2015 23:59
author: "{Your Name}"
output:
  html_document:
    toc: true 
---
# Network communities
Please send your reports to hse.ntwks@gmail.com with the subject of of the following structure:<br/>
*[MAGOLEGO SNA 2015] {LastName} {First Name} HA{Number}*

Late submission policy: -1 point per day

Use this file as a template for your report.\
Support your computations with figures and comments. Send ONLY .Rmd versions of your report.

```{r, echo=FALSE}
library(igraph)
```


## Yahoo Music Network

For this assignment, you have to load [the part of Yahoo Music Network](https://drive.google.com/file/d/0Bwd5a7NScgMtNElfOFdSbE1UYnM/view?usp=sharing). Archive contains network in GML format and corresponding list of artists.

```{r}
## Put your code here

```

Edges in this network appear if enough number of users have given ratings to both music bands. Note, that edges are weighted with similarity of the ratings.


### 1. Preprocessing

This network needs to be preprocessed. Do the steps, described on seminar, to make it look properly. How many vertices and edges were deleted?

```{r}
## Put your code here

```

### 2. Clustering

Define clusters for this networks using one of the algorithms described on lectures and seminars:
```{r}
## Put your code here

```

Compute density of obtained clusters:
```{r}
## Put your code here

```

Compute the ratio of inner clusters connections to outer ones:
```{r}
## Put your code here

```


### 3. Visualization & interpretation

Visulize five of the most dense clusters. Use names of artists as node labels on the graph.
```{r}
## Put your code here

```

<b> (Extra task) </b>Try to interpret (using Wikipedia or other resources) obtained clusters.

*Insert comments here*