---
title: "Social Network Analysis <br/> Home Assignment 3"
date: due date - 05.06.2015 23:59
author: "{Your Name}"
output:
  html_document:
    toc: true 
---
# Network communities
Please send your reports to hse.ntwks@gmail.com with the subject of of the following structure:<br/>
*[MAGOLEGO SNA 2015] {LastName} {First Name} HA{Number}*

Late submission policy: -1 point per day

Use this file as a template for your report.\
Support your computations with figures and comments. Send ONLY .Rmd versions of your report.

```{r, echo=FALSE}
library(igraph)
```


## Yahoo Music Network

For this assignment, you have to load [the part of Yahoo Music Network](https://drive.google.com/file/d/0Bwd5a7NScgMtNElfOFdSbE1UYnM/view?usp=sharing). Archive contains network in GML format and corresponding list of artists.

```{r}
## Put your code here
g <- read.graph(file = "music2K.gml", format = "gml")
art <- scan('artists.txt', character(0), sep = "\n")

for (cnt in 1:length(V(g)))
    V(g)$label[cnt] <- art[cnt]
```

Edges in this network appear if enough number of users have given ratings to both music bands. Note, that edges are weighted with similarity of the ratings.


### 1. Preprocessing

This network needs to be preprocessed. Do the steps, described on seminar, to make it look properly. How many vertices and edges were deleted?

```{r}
## Put your code here
vertex1 <- length(V(g))
edges1 <- length(E(g))
g <- delete.vertices(g, degree(g) == 0)
g <- simplify(g)
vertex2 <- length(V(g))
edges2 <- length(E(g))

vertex1 - vertex2
edges1 - edges2

```

### 2. Clustering

Define clusters for this networks using one of the algorithms described on lectures and seminars:
```{r}
## Put your code here
eb <- fastgreedy.community(g)
```

Compute density of obtained clusters:
```{r}
## Put your code here
elems <- list()
densities <- list()
for (i in 1:max(eb$membership)){
  g1 <- induced.subgraph(g, eb$membership == i)  
    
  e1 <- sum(eb$membership == i)
  d1 <- graph.density(g1)
  elems[i] <- e1
  densities[i] <- d1
  
  print(sprintf("Cluster # %d with %d elements has density = %.4f", i, e1, d1))
  }
```

Compute the ratio of inner clusters connections to outer ones:
```{r}
## Put your code here
for (i in 1:max(eb$membership)){
  h = induced.subgraph(g, eb$membership == i)
  edges_all = length(E(g))
  edges_in = length(E(h))
  edges_out = length(E(induced.subgraph(g,!(eb$membership == i))))
  edges_between = edges_all - edges_in - edges_out
  print(sprintf("Ratio for cluster # %d = %.3f",i, edges_in/edges_between))
}
```


### 3. Visualization & interpretation

Visulize three of the biggest clusters.
```{r}
## Put your code here
index_max1 <- which.max(elems)
elems[index_max1] <- 0
index_max2 <- which.max(elems)
elems[index_max2] <- 0
index_max3 <- which.max(elems)

subgraph <- induced.subgraph(g,(c((which(eb$membership == index_max1)),(which(eb$membership == index_max2)),((which(eb$membership == index_max3))))))
V(subgraph)[which(eb$membership == index_max1)]$color = 'black'
V(subgraph)[which(eb$membership == index_max2)]$color = 'blue'
V(subgraph)[which(eb$membership == index_max3)]$color = 'red'

plot(subgraph, layout=layout.fruchterman.reingold, vertex.size=3, edge.width = 0.005, vertex.label=NA)
```

Visulize three of the most dense clusters. Use artists' names as labels.
```{r}
index_max1 <- which.max(densities)
densities[index_max1] <- 0
index_max2 <- which.max(densities)
densities[index_max2] <- 0
index_max3 <- which.max(densities)

subgraph <- induced.subgraph(g,which(eb$membership == index_max1))
plot(subgraph, layout=layout.circle)
```


```{r}
subgraph <- induced.subgraph(g,which(eb$membership == index_max2))
plot(subgraph, layout=layout.circle)
```

```{r}
subgraph <- induced.subgraph(g,which(eb$membership == index_max3))
plot(subgraph, layout=layout.circle)
```

<b> (Extra task) </b>Try to interpret (using Wikipedia or other resources) obtained clusters.

*Insert comments here*