---
title: "Magolego SNA - Lab 5"
author: "Danil Gizdatullin, Evgenij Tsymbalov, Andrey Shestakov"
date: "05/15/2015"
output:
  html_document:
    toc: true 
---

```{r}
library('igraph')
```

**TO SAVE YOUR TIME, PLEASE START DOWNLOADING [THIS](https://www.cs.upc.edu/~csn/lab/wikipedia.gml) NETWORK RIGHT NOW **

## Cohesive subgraphs

### Graph cliques

Graph clique is a subset of vertices of a graph such that every two vertices  in the clique are adjacent.

How many cliques can you see on this graph?

```{r}
plot(graph.famous("bull"))
```

There was a couple of definitions about the cliques in graph on the lecture.

A <b>maximum clique</b> is a clique that cannot be extended by including one more adjacent vertex (not included in larger one). Can you name maximum cliques in the given graph?

A <b>maximal clique</b> is a clique of the largest possible size in a given graph. 

And, finally, <b>graph clique number</b> is the size of the maximum clique. Bull graph's clique number is 3.

`maximal.cliques` returns lists of vertices, that form a maximum graph.
Let's see maximum cliques for a bull graph:
```{r}
maximal.cliques(graph.famous("bull"))
```

Let's demonstrate some useful functions for finding cliques. Our graph today is again Zachary's Karate Club graph:

```{r}
g = graph.famous("Zachary")
plot(g)
```

We can define sizes of maximal cliques we interested in:
```{r}
maximal.cliques(g, min = 4, max = 5) # maximal cliques of sizes 4 and 5
```

`maximal.cliques` returns lists of vertices - maximal cliques.
`clique.number` returns graph's clique number.

Let's find and show maximal cliques for Zachary Carate Club graph:
`lrg = largest.cliques(g)` returns ids of nodes - largest cliques
```{r}

largest = largest.cliques(g)

op = par(mfrow = c(1,2))

labels = rep(0, vcount(g))

labels[largest[[1]]] = 2
plot(g, vertex.color = labels)
labels = rep(0, vcount(g))
labels[largest[[2]]] = 2
plot(g, vertex.color = labels)
par(op)
```

### k-core ###

**k-core** is a maximal subset of vertices such that each is connected to
at least k others in the subset.
  
R has a function wich calculates the *coreness* for each vertex. The coreness of a vertex is k if it belongs to the k-core but not to the (k+1)-core.

```{r}
# Let's make some graph
z<-graph.empty(n=11, directed = FALSE)
z <- add.edges(z,c(1,2, 1,3, 1,4, 1,6, 1,5, 2,3, 2,4, 3,10, 3,11, 3,8, 3,4, 4,8, 4,7, 8,9, 10,11))
plot(z)
```

Now we find maximum k-core and pick out it on graph

```{r}
coreness <- graph.coreness(z)
max_cor <- max(coreness)
max_cor
color_bar <- heat.colors(max_cor) 
plot(z, vertex.color = color_bar[coreness])
```

## Community detection

### The list of community detection algorithms in igraph

* `edge.betweenness.community` [Newman and Girvan, 2004]
* `fastgreedy.community` [Clauset et al., 2004] (modularity optimization method)
* `label.propagation.community` [Raghavan et al., 2007]
* `leading.eigenvector.community` [Newman, 2006]
* `multilevel.community` [Blondel et al., 2008] (the Louvain method)
* `optimal.community` [Brandes et al., 2008]
* `spinglass.community` [Reichardt and Bornholdt, 2006]
* `walktrap.community` [Pons and Latapy, 2005]
* `infomap.community` [Rosvall and Bergstrom, 2008]

### Newman-Girvan Edge-Betweenness ###

#### Edge betweenness ####

**Edge betweenness** is equal to the number of shortest paths from all vertices to all others that pass through that edge.

```{r}
g<-graph.empty(n=6, directed = FALSE)
g <- add.edges(g,c(1,2, 2,3, 1,3, 2,4, 4,5, 4,6, 5,6))
plot(g)
betw <- edge.betweenness(g)
#E(g)
#betw
```

#### The algorithm ####

The Newman-Girvan algorithm detects communities by progressively removing edges from the original network. The Girvan-Newman algorithm focuses on edges that are most likely "between" communities.
  
Algorithm:  
  
* <b>Step 1</b>: the betweenness of all existing edges in the network is calculated first.  
* <b>Step 2</b>: the edge with the highest betweenness is removed.  
* <b>Step 3</b>: the betweenness of all edges affected by the removal is recalculated.  
* <b>Step 4</b>: steps 2 and 3 are repeated until no edges remain.  

The best partition is selected based on modularity.

There is [edge.betweenness.community](http://www.inside-r.org/packages/cran/igraph/docs/edge.betweenness.community) function in R

```{r}
g <- graph.famous("Zachary")
eb <- edge.betweenness.community(g)
plot(eb, g)

## A bit more hand-made way
# color_map = c("grey","blue","black","yellow","red","green")
# membership = cutat(eb, no = 4)
# membership = eb$membership
# plot(g, vertex.color = eb$membership)

```

Also you can obtain dendrogram:
```{r}
dendPlot(eb, mode="hclust", rect = 5)

## Optionally you can run this
# dend <- as.dendrogram(eb)
# plot(dend)
```

### Greedy Modularity maximization ###

Alternatively to the previous method, this one is agglomerative. Intially consider a network s.t.
* There is no edges
* All clusters consist of a single vertex

Iteratively add an edge that delivers maximum modularity gain and merge correspondent communitues.

```{r}
g <- graph.famous(name = "Zachary")
mm <- fastgreedy.community(g)

plot(rev(mm$modularity), xlab = 'Number of clusters', ylab = 'Modularity value')
which.max(rev(mm$modularity))

plot(mm, g)
```

### Label propagation ###

Label propagation algorithm consists of four steps:

* <b>Step 1</b>: Initialize labels
* <b>Step 2</b>: Randomize node ordering
* <b>Step 3</b>: For every node replace its label with occurring with the highest frequency among neighbors
* <b>Step 4</b>: Repeat steps 2-3 until every node will have a label that the maximum number of its neighbors have

Warning! Due to **step 2** you may get different results.
```{r}
g <- graph.famous("Zachary")
lp <- label.propagation.community(g)
plot(lp, g)
```

### Wikipedia example ###

Load wikipedia network in R and run some community detection algorithm. Extract article names in some communities and check whether they make sense?

```{r}
g <- read.graph('wikipedia.gml', format = 'gml')
g <- as.undirected(g)
```

The next lines of code might be usefull for interpretation
```{r, warning=FALSE}
mm <- fastgreedy.community(g)
l <- V(g)$label[mm$membership == 2]
text <- paste(l, collapse = ' ')

#install.packages(c("tm", "SnowballC", "wordcloud", "RColorBrewer", "XML"))

library(wordcloud)

wordcloud(text, type="text", 
        lang="english", excludeWords = NULL, 
        textStemming = FALSE,  colorPalette="Dark2",
        max.words=200)
```