---
title: "Magolego SNA - Lab 6"
author: "Danil Gizdatullin, Evgenij Tsymbalov, Andrey Shestakov"
date: "05/15/2015"
output:
  html_document:
    toc: true 
---

## Motifs

<b>Motifs</b> are often defined as recurrent and statistically significant sub-graphs or patterns. Here, we consider motifs as sub-graphs of a given graph, which are isomorphic to defined sample.

Here is our old beloved friend:
```{r}
library('igraph')
g = graph.famous("Zachary")
```

Let's define a simple sample motif:
```{r}
sample = graph(c(c(1,2), c(2,3)))
plot(sample)
```

Now let's calculate, if it is often in our graph:
```{r}
isoclass_num = graph.isoclass(sample) # defining number of corresponding isoclass
motifs3 = graph.motifs(g, size = 3)
motifs3[isoclass_num] # returns number of motifs 
```

Due to high computational time (isomorphism checks), `graph.motifs` are implemented for graps of sizes 3 and 4 only. However, we can easy check numbers for all motifs of size 4 to find more frequent patterns:
```{r}
motifs4 = graph.motifs(g, size = 4)
```

The most frequent pattern is fifth. Let's draw it.
```{r}
plot(graph.isocreate(size=4, number=5))
```

Not what we're looking for.. Second most frequent (seventh):
```{r}
plot(graph.isocreate(size=3, number=7))
```

That is certainly better.

## Visualisation

There are currently three different functions in the igraph package which can draw graph in various ways:  

* `plot.igraph` does simple non-interactive 2D plotting to R devices. Actually it is an implementation of the plot generic function, so you can write plot(graph) instead of plot.igraph(graph). As it used the standard R devices it supports every output format for which R has an output device. The list is quite impressing: PostScript, PDF files, XFig files, SVG files, JPG, PNG and of course you can plot to the screen as well using the default devices, or the good-looking anti-aliased Cairo device. See [plot.igraph](http://igraph.org/r/doc/plot.graph.html) for some more information.  

* `tkplot` does interactive 2D plotting using the tcltk package. It can only handle graphs of moderate size, a thousend vertices is probably already too many. Some parameters of the plotted graph can be changed interactively after issuing the tkplot command: the position, color and size of the vertices and the color and width of the edges. [See tkplot](http://igraph.org/r/doc/tkplot.html) for details.  

* `rglplot` is an experimental function to draw graphs in 3D using OpenGL. See [rglplot](http://igraph.org/r/doc/rglplot.html) for some more information.  
  
Let's draw a graph-ring using different three methods.
```{r}
library(igraph)
library(rgl)

g <- graph.ring(10)
g$layout <- layout.circle
plot(g)

#doesn't work in R Markdown
tkplot(g)
rglplot(g)
```
###Layout###
Either a function or a numeric matrix. It specifies how the vertices will be placed on the plot.  
Let's demonstrate how it works on some graph. For example graph which based on barabashi model.

```{r}
g <- barabasi.game(50)
```

* `layout.auto` - tries to choose an appropriate layout function for the supplied graph, and uses that to generate the layout.  

```{r}
plot(g, layout=layout.auto, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```  

* `layout.random` - simply places the vertices randomly on a square.  

```{r}
plot(g, layout=layout.random, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```  

* `layout.circle` - places the vertices on a unit circle equidistantly. 

```{r}
plot(g, layout=layout.circle, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.sphere` - places the vertices (approximately) uniformly on the surface of a sphere, this is thus a 3d layout.  

```{r}
plot(g, layout=layout.sphere, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.fruchterman.reingold` uses a force-based algorithm proposed by Fruchterman and Reingold.  

```{r}
plot(g, layout=layout.fruchterman.reingold, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.kamada.kawai` is another force based algorithm.  

```{r}
plot(g, layout=layout.kamada.kawai, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.spring` is a spring embedder algorithm.  

```{r}
plot(g, layout=layout.spring, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.fruchterman.reingold.grid` is similar to layout.fruchterman.reingold but repelling force is calculated only between vertices that are closer to each other than a limit, so it is faster.  

```{r}
plot(g, layout=layout.fruchterman.reingold.grid, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.lgl` is for large connected graphs, it is similar to the layout generator of the Large Graph Layout software [http://lgl.sourceforge.net/](http://lgl.sourceforge.net/).  

```{r}
plot(g, layout=layout.lgl, vertex.size=4,
     vertex.label.dist=0.5, vertex.color="red", edge.arrow.size=0.5)
```

* `layout.graphopt` is a port of the graphopt layout algorithm by Michael Schmuhl. graphopt version 0.4.1 was rewritten in C and the support for layers was removed (might be added later) and a code was a bit reorganized to avoid some unneccessary steps is the node charge (see below) is zero.  

* `layout.svd` is a currently experimental layout function based on singular value decomposition.  

* `layout.norm` normalizes a layout, it linearly transforms each coordinate separately to fit into the given limits.  

* `layout.drl` is another force-driven layout generator, it is suitable for quite large graphs.  

* `layout.reingold.tilford` generates a tree-like layout, so it is mainly for tree.    
  
###Highlight components###
Let's make different color for each graph component  
  
```{r}
g <- erdos.renyi.game(100, 1/100)

l <- layout.fruchterman.reingold(g)

op = par(mfrow = c(1,2))
plot(g, layout=l, vertex.size=5, vertex.label=NA)

comps <- clusters(g)$membership
colbar <- rainbow(max(comps)+1)
V(g)$color <- colbar[comps+1]

plot(g, layout=l, vertex.size=5, vertex.label=NA)
```
  
###Highlight communities in graph###
Let's make different color for each community
  
```{r}
g <- graph.full(5) %du% graph.full(5) %du% graph.full(5)
g <- add.edges(g, c(1,6, 1,11, 6,11))

op = par(mfrow = c(1,2))
plot(g, layout = layout.kamada.kawai)

com <- spinglass.community(g, spins=5)
V(g)$color <- com$membership+1
g <- set.graph.attribute(g, "layout", layout.kamada.kawai(g))
plot(g, vertex.label.dist=1.5)
par(op)
```
  
###Trees Visualization###
```{r}
plot(graph.tree(50, 2))
```
We can use `layout = layout.reingold.tilford` to draw tree  

```{r}
plot(graph.tree(50, 2), vertex.size=3, vertex.label=NA, layout=layout.reingold.tilford)
tkplot(graph.tree(50, 2, mode="undirected"), vertex.size=10,
       vertex.color="green")
```

## Assortative Mixing

Assortative Mixing coefficient shows whether nodes with the same attribute values tend to form connections. Download [Caltech](https://www.dropbox.com/s/tz9teeaq79n0cx4/Caltech.gml?dl=0) friendship network. Inspect nodes attributes and compute assortativity coefficients with `assortativity` function

```{r}
# Your code here
g <- read.graph(file = 'Caltech.gml', format = 'gml')
assortativity.nominal(graph = g, types = V(g)$dorm+1, directed = F)

```

## Basic network analysis pipeline

The basic pipeline for exploratory graph analysis consists of:

### Loading
See Seminar 1.

### Cleaning
After loading graph with all necessary attributes, it is often recommended to:

* delete empty nodes:
```{r, eval = FALSE}
g = delete.vertices(g, degree(g) == 0)
```

* delete self-cycles and multiple edges - to make graph <i>simple</i>. `simplify` does all the work; some parameters could be tuned:
```{r}
simplify(g)
# is.simple(simplify(g, remove.multiple=FALSE))
```

###  Obtaining main characteristics
See Seminars 2, 4.

###  Clustering
See Seminar 5.

###  Visualization 
See previous section + Seminars 1, 4.
