---
title: "Social Network Analysis <br/> Home Assignment 4"
date: due date - 19.06.2015 23:59
author: "{Your Name}"
output:
  html_document:
    toc: true 
---
# Network Epidemics

Please send your reports to hse.ntwks@gmail.com with the subject of of the following structure:<br/>
*[MAGOLEGO SNA 2015] {LastName} {First Name} HA{Number}*

Late submission policy: -1 point per day

Use this file as a template for your report.\
Support your computations with figures and comments. Send ONLY .Rmd versions of your report.

```{r, echo=FALSE}
library(igraph)
```
## SIR Model

You need to perform epidemic SIR model on different types of networks:
Try different parameters for network generation
```{r}
gl <- list()
gl$ba <- barabasi.game(n = 100,m = 12 , directed=FALSE)
gl$er <- erdos.renyi.game(n = 250, p.or.m = 0.02, type=c("gnp"))
gl$ws <- watts.strogatz.game(dim = 1, size = 1000, nei = 3, p = 0.01)
``` 
Moreover perform modeling on real peer-to-peer network [here](https://www.dropbox.com/s/sto6dxyja0u7y4u/Net.txt?dl=0)


Your goal is to perform a research on epidemics:
Use different values of parameters listed below
```{r}
beta <- 2
gamma <- 4
ntrials <- 100

# Running function sir for each element of list gl (list with graphs)
sim <- lapply(gl, sir, beta=beta, gamma=gamma, no.sim=ntrials)
```
at least 3 different versions, for example:

* betta (4 6 8)
* gamma (8 6 2)
* niter (100 500 1000)
    
*For some reason `beta` and `gamma` parameters should not be set below 0 and 1. Looks like they are somehow normilized during simulation.*

The code below can help you with plotting
```{r}
plot(sim$er)
plot(sim$ba, color="palegoldenrod", median_color="gold", quantile_color="gold")
plot(sim$ws, color="pink", median_color="red", quantile_color="red")

x.max <- max(sapply(sapply(sim, time_bins), max))
y.max <- 1.05 * max(sapply(sapply(sim, function(x) median(x)[["NI"]]), max, na.rm=TRUE))

plot(time_bins(sim$er), median(sim$er)[["NI"]], type="l", lwd=2, col="blue", xlim=c(0, x.max), ylim=c(0, y.max), xlab="Time", ylab=expression(N[I](t)))
lines(time_bins(sim$ba), median(sim$ba)[["NI"]], lwd=2, col="gold")
lines(time_bins(sim$ws), median(sim$ws)[["NI"]],lwd=2, col="red")
legend("topright", c("ER", "BA", "WS"),col=c("blue", "gold", "red"), lty=1)
```

You need to plot three values on the graphics: Number of infected, number of suseprible, number of recovered - all depends on time.
As a result of this task, you need to provide 12 plots (one for each network with 3 diferent parameters) with explanation.


