---
title: "Social Network Analysis <br/> Home Assignment 2"
date: due date - 15.05.2015 23:59
author: "{Your Name}"
output:
  html_document:
    toc: true 
---
# Graph models. Centrality metrics
Please send your reports to hse.ntwks@gmail.com with the subject of of the following structure:<br/>
*[MAGOLEGO SNA 2015] {LastName} {First Name} HA{Number}*

Late submission policy: -1 point per day

Use this file as a template for your report.\
Support your computations with figures and comments. Send ONLY .Rmd versions of your report.

```{r, echo=FALSE}
library(igraph)
```


## Task 1. Your social network

For the first task, you have to load your [vk.com](http://vk.com) network. Please follow the instructions posted on the course wiki. 
If you did it correctly, you should have a GraphML file with your own network. Read it to R:

```{r}
## Put your code here

```

### 1. Degree distribution

First, plot degree distribution of your network in log-log scales:

```{r}
## Put your code here

```

Is there any correspondence between actual degree distribution of your network and the Power Law distribution? If not, explain why.

*Put your comments here*

Now, let's see how it would look if it was random. Produce Erdos-Renyi graph matching your real network (same number of nodes and same average degree). Compare it with your degree distribution.

```{r}
## Put your code here

```

*Put your comments here*

### 2. Compute centrality metrics

Compute for your network: 

* degree centrality

* closeness centrality

* betweenness centrality

* eigenvector centrality

* Bonacich power centrality

* Alpha centrality


```{r}
## Put your code here

```

Output six plots corresponding to six centrality metrics you've computed: 

* Use first names of your friends as node labels on the graph (you may hide this information if you wish -- change it by integer ID)
* Keep the same layout of the network
* Make node sizes and colours proportional to the respective centrality metrics

Hint: If you get stuck, lab 4 script might be helpful.

```{r}
## Put your code here

```

Now, output top ten nodes in each ranking. Again, print only first names in your table to keep privacy:

```{r}
## Put your code here

```

*Comment on your results here - for example, why some nodes have high betweenness centrality and the others have high closeness centrality. Is this what you would expect to see?*


## Task 2. Flickr network

In the second task, you will work with a large directed graph. 

Please download [flickr.mat](https://www.dropbox.com/s/the9wqk987oxbpf/flickr.mat?dl=0) 

*Data contains sparse matrix A and list of user names. This is a denser part of the Flickr photo sharing site friendship graph from 2006. Edge direction corresponds to friendship requests (following). Some of the links are reciprocal, others not*

It's a Matlab file. How to deal with it in R? There is a package `R.matlab`. Please install it and call `library(R.matlab)`

Now use `readMat` function to read the file and extract adjacency matrix and a list of user names:

```{r, eval=FALSE}
flickr = readMat("YOUR_PATH/flickr.mat")
fmatrix=as.matrix(flickr[1]$A)
fnames=flickr[2]$names
```

Look at user names. You might want to remove spaces from the names. Use a function `gsub` to remove them:

```{r, eval=FALSE}
fnames=gsub(" ", "", fnames)
```

Now create a graph, output the number of vertices and edges:

```{r}
## Put your code here

```

Compute in- and out- degree centralities, PageRank, Hubs and Authorities for this network:

```{r}
## Put your code here

```

Print top ten names in each ranking:

```{r}
## Put your code here

```

Produce the following plots:

* In-degree centralities versus out-degree centralities

* In-degree centralities versus authorities

* Out-degree centralities versus hubs

* Hubs versus authorities

* PageRank versus hubs

* PageRank versus authorities


```{r}
## Put your code here

```

*Comment on the relationships between different centrality metrics*
